Large squeezing in multicomponent trapped Bose-Einstein condensates

### 2013-06-11 large_number_squeezing_with_atoms.tex
Complete the description of the scaling laws.

Good squeezing requires that the Bogoliubov fraction or condensate depletion be small.  To better understand the scaling of $N_{\mathbf{k}\neq0}/N$, we will approximate the sum in \eqref{eqCondensateDepletion} with an integral.  In 1D \eqref{eqCondensateDepletion} becomes
\begin{align}
  \frac{N_{\mathbf{k}\neq 0}(t)}{N} &= \frac{(n_a \chi_{aa} \tau_\text{hold} \cos\theta)^2}{N} \sum_{k \neq 0} \sinc^2\left(\sqrt{\omega^0_{k}(\omega^0_{k}+2 \chi_{aa}n_a)}\tau_\text{hold}\right).
\end{align}
In the limit that $\omega_{k_\text{min}} \tau_\text{hold} \ll 1$, where $k_\text{min} = 2\pi/L$ this discrete sum may be approximated by the integral
\begin{align}
  \frac{N_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{(n_a \chi_{aa} \tau_\text{hold} \cos\theta)^2}{N} \int_{-\infty}^{+\infty} \sinc^2 \left(\sqrt{\omega^0_{k}(\omega^0_{k}+2 \chi_{aa}n_a)}\tau_\text{hold}\right) \frac{d k}{\Delta k},
\end{align}
where $\Delta k = 2\pi/L$.  Changing integration variables to $\Phi = \omega^{0}_{\mathbf{k}}\tau_\text{hold}$, the integral can be written as
\begin{align}
  \frac{N_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{(n_a \chi_{aa} \tau_\text{hold} \cos\theta)^2}{N} \frac{L}{\sqrt{2}\pi\sqrt{\frac{\hbar}{M}\tau_\text{hold}}} \int_0^{\infty} \frac{1}{\sqrt{\Phi}}\sinc^2\sqrt{\Phi(\Phi + 2 \lambda_{aa} n_a)}\, d\Phi
\end{align}
Define $\Lambda_{a} = n_a \chi_{aa} \tau_\text{hold}$ to simplify the expression and we get
\begin{align}
  \frac{N^\text{1D}_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{1}{N} \frac{\Lambda_a^2 L \cos^2\theta } {\sqrt{2}\pi\sqrt{\frac{\hbar}{M}\tau_\text{hold}}} f_\text{1D}(\Lambda_a),
\end{align}
where
\begin{align}
  f_\text{1D}(\Lambda) &= \int_0^{\infty} \frac{1}{\sqrt{\Phi}}\sinc^2\sqrt{\Phi(\Phi + 2 \Lambda)}\, d\Phi
\end{align}

In 2D and 3D you get
\begin{align}
  \frac{N^{\text{2D}}_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{1}{N} \frac{\Lambda_a^2 A \cos^2\theta } {4\frac{\hbar}{M}\tau_\text{hold}} f_\text{2D}(\Lambda_a), \\
  \frac{N^{\text{3D}}_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{1}{N} \frac{\Lambda_a^2 V \cos^2\theta } {\sqrt{2}\pi^2\left(\frac{\hbar}{M}\tau_\text{hold}\right)^{3/2}} f_\text{3D}(\Lambda_a),
\end{align}
where
\begin{align}
  f_\text{2D}(\Lambda) &= \int_0^{\infty} \sinc^2\sqrt{\Phi(\Phi + 2 \Lambda)}\, d\Phi\\
  f_\text{3D}(\Lambda) &= \int_0^{\infty} \sqrt{\Phi} \sinc^2\sqrt{\Phi(\Phi + 2 \Lambda)}\, d\Phi
\end{align}

>>>>

Good squeezing requires that the Bogoliubov fraction or condensate depletion be small.  To better understand the scaling of $N_{\mathbf{k}\neq0}/N$, we will approximate the sum in \eqref{eqCondensateDepletion} with an integral.  In 1D \eqref{eqCondensateDepletion} becomes
\begin{align}
  \frac{N_{\mathbf{k}\neq 0}(t)}{N} &= \frac{(n_a \chi_{aa} \tau_\text{hold} \cos\theta)^2}{N} \sum_{k \neq 0} \sinc^2\left(\sqrt{\omega^0_{k}(\omega^0_{k}+2 \chi_{aa}n_a)}\tau_\text{hold}\right).
\end{align}
In the limit that $\omega_{k_\text{min}} \tau_\text{hold} \ll 1$, where $k_\text{min} = 2\pi/L$ this discrete sum may be approximated by the integral
\begin{align}
  \frac{N_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{(n_a \chi_{aa} \tau_\text{hold} \cos\theta)^2}{N} \int_{-\infty}^{+\infty} \sinc^2 \left(\sqrt{\omega^0_{k}(\omega^0_{k}+2 \chi_{aa}n_a)}\tau_\text{hold}\right) \frac{d k}{\Delta k},
\end{align}
where $\Delta k = 2\pi/L$.  Changing integration variables to $\Phi = \omega^{0}_{\mathbf{k}}\tau_\text{hold}$, the integral can be written as
\begin{align}
  \frac{N_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{(n_a \chi_{aa} \tau_\text{hold} \cos\theta)^2}{N} \frac{L}{\sqrt{2}\pi\sqrt{\frac{\hbar}{M}\tau_\text{hold}}} \int_0^{\infty} \frac{1}{\sqrt{\Phi}}\sinc^2\sqrt{\Phi(\Phi + 2 \lambda_{aa} n_a)}\, d\Phi
\end{align}
Define $\Lambda_{a} = n_a \chi_{aa} \tau_\text{hold}$ to simplify the expression and we get
\begin{align}
  \frac{N^\text{1D}_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{1}{N} \frac{\Lambda_a^2 L \cos^2\theta } {\sqrt{2}\pi\sqrt{\frac{\hbar}{M}\tau_\text{hold}}} f^\text{1D}(\Lambda_a),
\end{align}
where
\begin{align}
  f^\text{1D}(\Lambda) &= \int_0^{\infty} \frac{1}{\sqrt{\Phi}}\sinc^2\sqrt{\Phi(\Phi + 2 \Lambda)}\, d\Phi
\end{align}

In 2D and 3D you get
\begin{align}
  \frac{N^{\text{2D}}_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{1}{N} \frac{\Lambda_a^2 A \cos^2\theta } {4\frac{\hbar}{M}\tau_\text{hold}} f^\text{2D}(\Lambda_a), \\
  \frac{N^{\text{3D}}_{\mathbf{k}\neq 0}(t)}{N} &\approx \frac{1}{N} \frac{\Lambda_a^2 V \cos^2\theta } {\sqrt{2}\pi^2\left(\frac{\hbar}{M}\tau_\text{hold}\right)^{3/2}} f^\text{3D}(\Lambda_a),
\end{align}
where
\begin{align}
  f^\text{2D}(\Lambda) &= \int_0^{\infty} \sinc^2\sqrt{\Phi(\Phi + 2 \Lambda)}\, d\Phi\\
  f^\text{3D}(\Lambda) &= \int_0^{\infty} \sqrt{\Phi} \sinc^2\sqrt{\Phi(\Phi + 2 \Lambda)}\, d\Phi
\end{align}

The squeezing predicted by the two-mode model will only be valid in higher dimensions when $N_{\mathbf{k}\neq0}/N \ll 1$, when this condition is violated, the squeezing will be detrimentally affected.  We therefore want to examine the scaling of $N_{\mathbf{k}\neq 0}/N$ with system size in 1D, 2D and 3D.

For constant total number $N$ and occupations of the $\ket{a}$ and $\ket{b}$ states at time $t_1$, the optimum squeezing as determined by the two-mode model will have a particular value of $\tau_\text{hold}$, $\theta$ and $\phi$.  As the dimensions of the system are scaled, $\hbar \chi_{ij} = U^\text{1D}_{ij}/L = U^\text{2D}_{ij}/A = U^\text{3D}_{ij}/V$ will change, but to keep the same squeezing (and the same equivalent two-mode model) $\lambda_{ij} = \chi_{ij} \tau_\text{hold}$ must be kept constant (see Eq.~\eqref{eqsDef}--\eqref{eqNumDiffVariance}).  Therefore $\Lambda_a = n_a\chi_{aa}\tau_\text{hold}$ must also be kept constant. With this in mind, we can write $\tau_\text{hold}$ in terms of the system size and the nonlinearity.  We obtain:
\begin{align}
  \frac{N^\text{1D}_{\mathbf{k}\neq 0}(t_3)}{N} &\approx \sqrt{L} \cdot \frac{1}{N} \sqrt{\frac{M U^\text{1D}_{aa} n_a}{2 \hbar^2 \pi^2}} \Lambda_a^{3/2} \cos^2\theta f^\text{1D}(\Lambda_a) &&\propto \sqrt{L} \\
  \frac{N^\text{2D}_{\mathbf{k}\neq 0}(t_3)}{N} &\approx 1 \cdot \frac{1}{N} \frac{M U^\text{2D}_{aa} n_a }{4 \hbar^2} \Lambda_a \cos^2\theta f^\text{2D}(\Lambda_a) &&\propto \text{constant} \\
  \frac{N^\text{3D}_{\mathbf{k}\neq 0}(t_3)}{N} &\approx \frac{1}{\sqrt{V}} \cdot \frac{1}{N} \left(\frac{M U^\text{3D}_{aa} n_a}{\hbar^2}\right)^{3/2} \frac{\Lambda_a^{1/2} \cos^2\theta f^\text{3D}(\Lambda_a)}{\sqrt{2} \pi^2} &&\propto \frac{1}{\sqrt{V}}
\end{align}

In 1D the condensate depletion $N_{\mathbf{k}\neq 0}/N$ increases with increasing system size, in 2D the condensate depletion remains constant, and in 3D the condensate depletion decreases as the system size increases.  This suggests that the best system to use will be a box potential, and the system size can be increased until the condensate depletion becomes negligibly small.
